{RandomRationalPoints} package for \emph{Macaulay2}

### 2021-01-21 FastLinAlg/Paper/RandomRationalPoints.tex
Minor Corrections

    There is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one point quickly, even if it is not rational.  
    We develop functions that apply different strategies to generate random rational and geometric points onn the given variety. We also provide functions that will expedite the process of determining properties of the singular locus of $X$.% when combined with the functions provided in the package {\tt FastLinAlg}.

@@

\section{Our primary pupose: {\tt randomPoints}}\label{randomPoints}
    We start with the core function in this package:  {\tt randomPoints} is a function to find rational or geometric points in a variety. The typical usages are as follows:

@@

This function takes a polynomial ring as an input and outputs the coordinate change map.  
i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.

@@

Alternately, instead of {$I$}, we may pass it a quotient ring.  It will then return the inclusion of the generic projection ring into the given ring, followed by the source of that inclusion.  It is essentially the same functionality as calling {\tt genericProjection(n, ideal R)} although the format of the output is slightly different.

@@

\vspace{0.5em}
-- {\tt findANonZeroMinor(n, M, I)} 
\vspace{0.5em}
\noindent where $I$ is an ideal
in a polynomial ring over {\tt QQ} or {\tt ZZ/p} for $p$ prime; $M$ is a matrix
over the polynomial ring and $n\in \mathbb{Z}$ denotes the size of the minors of interest.

@@

\item[\tt MinorPointAttempts => ZZ] 
  This controls how many points at which to check the rank of the matrix.
\end{description}
\vspace{0.25em}
\begin{example}
  We demonstrate this function.
  ~~
  {{\small\color{blue}
  \begin{quote}
\begin{verbatim}
		i3 : R = ZZ/5[x, y, z];

@@

		o6 : Sequence
\end{verbatim}
\end{quote}\vspace{-1em}    
    }}
\end{example}

@@

\end{verbatim}
\end{quote}\vspace{-1em}    
    }}    
    \noindent
    In this particular example, there tend to be about $5$ associated primes when adding the first minor to $J$, and so one would expect about $5$ steps as each minor computed most likely will eliminate one of those primes.
\end{example}

There is some similar functionality obtained via heuristics (as opposed to actually finding rational points) in the package {\tt FastLinAlg}.

\vspace{2em}
\end{document}

>>>>

    There is an existing package called {\tt RandomPoints}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one point quickly, even if it is not rational.  
    We develop functions that apply different strategies to generate random rational and geometric points on the given variety. We also provide functions that will expedite the process of determining properties of the singular locus of $X$.% when combined with the functions provided in the package {\tt FastLinAlg}.

@@

\section{Our primary purpose: {\tt randomPoints}}\label{randomPoints}
    We start with the core function in this package:  {\tt randomPoints} is a function to find rational or geometric points in a variety. The typical usages are as follows:

@@

This function takes a polynomial ring as an input and outputs the coordinate change map,
i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.

@@

Alternatively, instead of {$I$}, we may pass it a quotient ring.  It will then return the inclusion of the generic projection ring into the given ring, followed by the source of that inclusion.  It is essentially the same functionality as calling {\tt genericProjection(n, ideal R)} although the format of the output is slightly different.

@@

\vspace{0.5em}
-- {\tt findANonZeroMinor(n, M, I)} 
\vspace{0.5em}

\noindent where $I$ is an ideal
in a polynomial ring over {\tt QQ} or {\tt ZZ/p} for $p$ prime, $M$ is a matrix
over the polynomial ring and $n\in \mathbb{Z}$ denotes the size of the minors of interest.

@@

\item[\tt MinorPointAttempts => ZZ] 
  This controls how many points at which to check the rank of the matrix.
\end{description}
%\vspace{0.25em}
\begin{example}
  We demonstrate this function.
  {{\small\color{blue}
  \begin{quote}
\begin{verbatim}
		i3 : R = ZZ/5[x, y, z];

@@

		o6 : Sequence
\end{verbatim}
\end{quote}    
    }}
\end{example}

@@

\end{verbatim}
\end{quote}
    }}    
    \noindent
    In this particular example, there tend to be about $5$ associated primes when adding the first minor to $J$, and so one would expect about $5$ steps as each minor computed most likely will eliminate one of those primes.
\end{example} 
There is some similar functionality obtained via heuristics (as opposed to actually finding rational points) in the package {\tt FastLinAlg}.

% \end{enumerate}
\vspace{1em}
\end{document}
